---
title: "term paper"
author: "Huang Po-Jui"
date: "2023-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# load game data
```{r}
game<- read.csv("C:/Users/user0522/Documents/GitHub/Baseball_db_HPJ/unstandardized_data.csv")
```

# data pre-process
### check na
```{r}
colSums(is.na(game))
```

### replace na with mean value
```{r}
game$release_spin_rate[is.na(game$release_spin_rate)] <- mean(game$release_spin_rate, na.rm = TRUE)
game$release_extension[is.na(game$release_extension)] <- mean(game$release_extension, na.rm = TRUE)
game$spin_axis[is.na(game$spin_axis)] <- mean(game$spin_axis, na.rm = TRUE)
game$BB.[is.na(game$BB.)] <- mean(game$BB., na.rm = TRUE)
```


### check na again
```{r}
colSums(is.na(game))
```


### check column type
```{r}
str(game)
```

# Feature Selection
### select X and Y
```{r}
library(dplyr)

df = select(game, -c('X', 'description', 'stand', 'p_throws', 'inning_topbot', 'Name', 'Pitcher'))
```


### check dimension
```{r}
dim(df)
```


### LASSO feature selection
```{r}
library(glmnet)
library(caret)
library(plotmo)

library(Matrix)
library(ggplot2)
library(lattice)
library(Formula)
library(plotrix)
library(TeachingDemos)

control = trainControl(method = "cv", number = 10) 

set.seed(123)

LASSO_fit <- train(result ~ .-1, data = df, method = "glmnet",    
                trControl = control,
                tuneGrid = expand.grid(alpha = 1, lambda = 0))

```


```{r}
coef(LASSO_fit$finalModel, LASSO_fit$bestTune$lambda)
```


```{r}
X_lasso = model.matrix(result ~ .-1, data=df)
Y_lasso = df$result

df_LASSO = glmnet(X_lasso, Y_lasso, alpha = 1)

print(df_LASSO)
```

```{r}
coef(df_LASSO, s = cv.glmnet(X_lasso, Y_lasso)$lambda.1se)
```

### finding best lambda
```{r}
cv.glmnet(X_lasso, Y_lasso)$lambda.1se
```


```{r}
plot(df_LASSO, label = TRUE)
```

**as lambda increasing, the coefficient of feature mostly goes to 0**
```{r}
par(mfrow=c(1,2))

plot_glmnet(df_LASSO, xvar = "lambda", 
            label = 5,   # the "label" means the top-n variable you want the graph to show (we want it to display the top-5 variables)
             xlab = expression(paste("log(", lambda, ")")),   # use expression() to ask R to enable Greek letters typesetting
             ylab = expression(beta))  
abline(v = log(cv.glmnet(X_lasso, Y_lasso)$lambda.1se), lty = "dashed") 

plot_glmnet(df_LASSO, label = 5, xlab = expression(paste("log(", lambda, ")")), ylab = expression(beta))
abline(v = log(cv.glmnet(X_lasso, Y_lasso)$lambda.1se), lty = "dashed")

## as lambda increasing, the coefficient of feature mostly goes to 0
```

#### important features of LASSO:<br> 
outs_when_up -0.004144210<br> 
delta_run_exp -2.142785769<br> 
pitch_type_CU 0.138911130<br> 
zone_4.0 -0.003590574<br> 
zone_5.0 -0.009671238<br> 
zone_6.0 -0.004523019<br> 
zone_8.0 -0.009330144<br>
zone_11.0 0.418600911<br>
zone_12.0 0.441588841<br>
zone_13.0 0.419897914<br>
zone_14.0 0.403122251<br>

# using feature seleted df to fit model
### create feature selected df 
```{r}
lasso_selected_df = select(game, c("result", "outs_when_up", "delta_run_exp", "pitch_type_CU", "zone_4.0", "zone_5.0", "zone_6.0", "zone_8.0", "zone_11.0", "zone_12.0", "zone_13.0", "zone_14.0"))
```


```{r}
library(e1071)
library(caret)
library(gbm)
library(pROC)
library(plotROC)
library(ISLR)
```


### train test split
```{r}
set.seed(123)

s_train_row_number <- createDataPartition(lasso_selected_df$result, p=0.75, list=FALSE)

s_train <- lasso_selected_df[s_train_row_number, ]
s_test <- lasso_selected_df[-s_train_row_number, ]
```


**logit**
```{r}
s_logit_model <- glm(result ~ ., family = binomial, data = s_train)
```


**probit**
```{r}
s_probit_model <- glm(result ~ ., family = binomial(link = "probit"), data = s_train)
```


**GBM**
```{r}
library(rsample) 
library(gbm)
library(xgboost)
library(pdp)
library(ggplot2)
library(purrr)
```


```{r}
hyper_grid <- expand.grid(
 shrinkage = c(.01, .1, .2),
 interaction.depth = c(1, 3, 5),
 n.minobsinnode = c(5, 10, 15),
 bag.fraction = c(.5, .75, 1),
 optimal_trees = 0, # you will fill in values from loop
 min_RMSE = 0) 
```


```{r}
for(i in 1:nrow(hyper_grid)) {
  
  # reproducibility
  set.seed(123)
  
  # train model
  gbm.tune <- gbm(
    result ~ .,
    data = s_train,
    distribution = "bernoulli",
    cv.folds = 10,
    n.trees = 500,
    interaction.depth = hyper_grid$interaction.depth[i],
    shrinkage = hyper_grid$shrinkage[i],
    n.minobsinnode = hyper_grid$n.minobsinnode[i],
    bag.fraction = hyper_grid$bag.fraction[i],
    train.fraction = .75,
    n.cores = NULL, # will use all cores by default
    verbose = FALSE
  )
  
  # locate minimum training error from the n-th tree, add it to grid
  hyper_grid$optimal_trees[i] <- which.min(gbm.tune$valid.error)
  hyper_grid$min_RMSE[i] <- sqrt(min(gbm.tune$valid.error))
}
```


```{r}
optimal_para <- hyper_grid %>% dplyr::arrange(min_RMSE) %>% head(1)

# define a list of optimal hyperparameters
shrink <- optimal_para$shrinkage
ntrees <- optimal_para$optimal_trees
minobs <- optimal_para$n.minobsinnode
bagfrac <- optimal_para$bag.fraction
depth <- optimal_para$interaction.depth

# train final GBM model
s_gbm_model <- gbm(
  result ~ .,
  data = s_train,
  distribution = "bernoulli",
  cv.folds = 10,
  n.trees = ntrees,
  interaction.depth = depth,
  shrinkage = shrink,
  n.minobsinnode = minobs,
  bag.fraction = bagfrac, 
  n.cores = NULL, # will use all cores by default
  verbose = FALSE
  )  
```


```{r}
print(shrink) 
print(ntrees) 
print(minobs) 
print(bagfrac) 
print(depth)
```


**svm**
```{r}
lasso_selected_df_for_svm = lasso_selected_df

svm_train <- lasso_selected_df_for_svm[s_train_row_number, ]
svm_test <- lasso_selected_df_for_svm[-s_train_row_number, ]

svm_train$result <- as.factor(svm_train$result)
svm_test$result <- as.factor(svm_test$result)

s_svm_model <- svm(result ~ ., 
                 scale = TRUE, # the default, x and y are scaled to zero mean and unit variance
               kernel = "radial", # other options are available
               degree = 3, # if kernel is of type = "polynomial"
                # gamma = if (is.vector(x)) 1 else 1 / ncol(x), # you can provide one for variable x
               cost = 1, # the cost function (C)
               probability = FALSE, # whether to output probability predictions
               na.action = na.omit,
                 data = svm_train)
```



**tree**
```{r}
library(rsample)
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)
```


```{r}
lasso_selected_df_for_tree = lasso_selected_df

lasso_selected_df_for_tree$result <- as.factor(lasso_selected_df_for_tree$result)

s_train_tree <- lasso_selected_df_for_tree[s_train_row_number, ]
s_test_tree <- lasso_selected_df_for_tree[-s_train_row_number, ]



hyper_grid <- expand.grid(
  minsplit = seq(5, 20, 1),  # from 5 to 20, with an increment of 1
  maxdepth = seq(8, 15, 1)
)

trees <- list()

for (i in 1:nrow(hyper_grid)) {
  
  # get minsplit, maxdepth values at row i
  minsplit <- hyper_grid$minsplit[i]
  maxdepth <- hyper_grid$maxdepth[i]
  # train a model and store in the list
  # R will feed i-th minsplit and maxdepth values from hyper_grid into rpart()'s control function as hyperparameters
  trees[[i]] <- rpart(
    result ~ .,
    data = s_train_tree,
    method  = "anova",
    control = list(minsplit = minsplit, maxdepth = maxdepth)
    )
}

get_cp <- function(x) {
  min <- which.min(x$cptable[, "CP"]) # to select from the column named "CP" and return a row index
  cp <- x$cptable[min, "CP"]  # using the index to subset the desired value
}

get_min_error <- function(x) {
  min    <- which.min(x$cptable[, "xerror"]) # to select from the column named "xerror" and return an index
  xerror <- x$cptable[min, "xerror"]  # using the index to subset the desired value
}

hyper_grid <- hyper_grid %>%
  mutate(
    cp    = purrr::map_dbl(trees, get_cp),  # purrr is a useful functional programming command from the tidyverse package, map_dbl() outputs double vectors (i.e, numeric values that have decimals) from an element of an object; mutate() adds new variables while preserving existing ones (from a data.frame or data.table object)
    error = purrr::map_dbl(trees, get_min_error)
    ) %>%
  arrange(error) %>%
  top_n(-5, wt = error)

min_split <- hyper_grid[1, 1]
max_depth <- hyper_grid[1, 2]
complexity_para <- hyper_grid[1, 3]

optimal_tree <- rpart(
    result ~ .,
    data = s_train_tree,
    # method  = "anova", in order to get predicted class, we need to suppress the "anove" method
    control = list(minsplit = min_split, maxdepth = max_depth, cp = complexity_para)
    )
```

```{r}
print(min_split)
print(max_depth)
print(complexity_para)
```



```{r}
head(trees, 1)
```

**RF**
```{r}
library(randomForest)
library(rsample)     # data splitting 
library(dplyr)       # data wrangling
library(rpart)       # build tree models
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)  
library(ranger)
library(ggplot2)
```


```{r}
set.seed(123)

hyper_grid_2 <- expand.grid(
  mtry = seq(1, 4, by = 1),
  node_size = seq(3, 9, by = 2),
  sampe_size = c(.5, .6, .75, .80),
  OOB_RMSE = 0
)

for(i in 1:nrow(hyper_grid_2)) {
  
  # train model
  model <- ranger(
    result ~ ., 
    data = s_train_tree, 
    num.trees = 100,
    mtry = hyper_grid_2$mtry[i],
    min.node.size = hyper_grid_2$node_size[i],
    sample.fraction = hyper_grid_2$sampe_size[i],
    seed = 123
  )
  
  # add OOB error to grid. note how we append the list of values to the vector and index them (fill in and replace the 0 inside the OOB_RMSE column)
  hyper_grid_2$OOB_RMSE[i] <- sqrt(model$prediction.error)
}

best_para <- hyper_grid_2 %>% 
  dplyr::arrange(OOB_RMSE) %>%
  head(1)

m <- best_para$mtry
node_size <- best_para$node_size
p <- best_para$sampe_size


optimal_ranger <- ranger(
    result ~ ., 
    data = s_train_tree, 
    num.trees = 100,
    mtry = m, # number of features
    min.node.size = node_size, # node size
    sample.fraction = p, # fraction of sample used in each tree
    importance = 'impurity'
  )
```


```{r}
print(m)
print(node_size)
print(p)
print(best_para)
```



# prediction
**logit model**
```{r}
pred_logit <- predict(s_logit_model, s_test, type = "response")

# classify outcomes by threshold = 0.5
convert_logit <- as.numeric(pred_logit > 0.5)

# generate confusion matrix and calculate precision score
confmat_logit <- table(Predicted = convert_logit, Actual = s_test$result)

confmat_logit
```


```{r}
s_logit_model$coefficients
```


```{r}
accuracy_logit <- (confmat_logit[1, 1] + confmat_logit[2, 2]) / sum(confmat_logit)
print(paste0("Accuracy of Logit Model: ",accuracy_logit))
```



**probit model**
```{r}
pred_probit <- predict(s_probit_model, s_test, type = "response")

convert_probit <- as.numeric(pred_probit > 0.5)

confmat_probit <- table(Predicted = convert_probit, Actual = s_test$result)

confmat_probit
```


```{r}
s_probit_model$coefficients
```


```{r}
accuracy_probit <- (confmat_probit[1, 1] + confmat_probit[2, 2]) / sum(confmat_probit)
print(paste0("Accuracy of Probit Model: ", accuracy_probit))
```


**svm**
```{r}
pred_svm <- predict(s_svm_model, s_test)

confmat_svm <- table(Predicted = pred_svm, Actual = s_test$result)

confmat_svm
```


```{r}
accuracy_svm <- (confmat_svm[1, 1] + confmat_svm[2, 2]) / sum(confmat_svm)
print(paste0("Accuracy of SVM Model: ",accuracy_svm))
```



**GBM**
```{r}
pred_gbm <- predict(s_gbm_model, s_test, type = "response")

convert_gbm <- as.numeric(pred_gbm > 0.5)

confmat_gbm <- table(Predicted = convert_gbm, Actual = s_test$result)

confmat_gbm
```


```{r}
accuracy_gbm <- (confmat_gbm[1, 1] + confmat_gbm[2, 2]) / sum(confmat_gbm)
print(paste0("Accuracy of GBM Model: ",accuracy_gbm))
```


##### relative influence of GBM model
```{r}
par(mar = c(5, 8, 1, 1)) 

summary(
  s_gbm_model, 
  cBars = 5,  # since we have 5 variables to display
  method = relative.influence, # you can also use permutation.test.gbm
  las = 2
  )
```



**decision tree**
```{r}
# predict on testing data
pred_tree <- predict(optimal_tree, newdata = s_test_tree, type = "class") # set type to "class"

# create confusion matrix and generate accuracy score
confmat_tree <- table(Predicted = pred_tree, Actual = s_test_tree$result)

# view the confusion table (lots of Type II error)
confmat_tree
```


```{r}
accuracy_tree <- (confmat_tree[1, 1] + confmat_tree[2, 2]) / sum(confmat_tree)
print(paste0("Accuracy of Tree Model: ",accuracy_tree))
```


##### plot tree model
```{r}
rpart.plot(optimal_tree)
```


**RF**
```{r}
# predict on testing data
pred_rf <- predict(optimal_ranger, s_test_tree, type = "response") # set type to "class"

# create confusion matrix and generate accuracy score
confmat_rf <- table(Predicted = pred_rf$predictions, Actual = s_test$result)

# view the confusion table (lots of Type II error)
confmat_rf
```


```{r}
accuracy_rf <- (confmat_rf[1, 1] + confmat_rf[2, 2]) / sum(confmat_rf)
print(paste0("Accuracy of Random Forest Model: ",accuracy_rf))
```


##### feature importanct of random forest
```{r}
rf_feature_importance <- data.frame(vname = names(lasso_selected_df[-1]), importance = optimal_ranger$variable.importance)  

ggplot(rf_feature_importance, aes(x = reorder(vname, -importance), y = importance, fill = vname, label = round(importance, digits = 2))) +  # fill means label with variable names
  geom_bar(stat = "Identity") +
  ggtitle("Most important variables (in the order of 1 to 4)") +
  xlab("Variable") + 
  ylab("Importance") +
  geom_text(size = 3, position = position_stack(vjust = 0.5), col = "white")
```


# comparison: svm model without lasso feature selection
```{r}
compare_df_for_svm = df

compare_svm_train <- compare_df_for_svm[s_train_row_number, ]
compare_svm_test <- compare_df_for_svm[-s_train_row_number, ]

compare_svm_train$result <- as.factor(compare_svm_train$result)
compare_svm_test$result <- as.factor(compare_svm_test$result)

compare_svm_model <- svm(result ~ ., 
                 scale = TRUE, # the default, x and y are scaled to zero mean and unit variance
               kernel = "radial", # other options are available
               degree = 3, # if kernel is of type = "polynomial"
                # gamma = if (is.vector(x)) 1 else 1 / ncol(x), # you can provide one for variable x
               cost = 1, # the cost function (C)
               probability = FALSE, # whether to output probability predictions
               na.action = na.omit,
                 data = compare_svm_train)

pred_svm_compare <- predict(compare_svm_model, compare_svm_test)

confmat_svm_compare <- table(Predicted = pred_svm_compare, Actual = compare_svm_test$result)

confmat_svm_compare

```


```{r}
accuracy_compare_svm <- (confmat_svm_compare[1,1] + confmat_svm_compare[2,2])/sum(confmat_svm_compare)
print(paste0("Accuracy of Comparision SVM Model:", accuracy_compare_svm))
```



```{r}
svm.pred2 <- as.numeric(pred_svm_compare)
ROC_svm_compare <- roc(s_test$result, svm.pred2, percent = TRUE, main = "Smoothing")
auc(ROC_svm_compare)
```


# ROC & AOU
```{r}
ROC_logit <- roc(s_test$result, pred_logit, percent = TRUE, main = "Smoothing")
ROC_probit <- roc(s_test$result, pred_probit, percent = TRUE, main = "Smoothing")

svm.pred1 <- as.numeric(pred_svm)
ROC_svm <- roc(s_test$result, svm.pred1, percent = TRUE, main = "Smoothing")
ROC_gbm <- roc(s_test$result, pred_gbm, percent = TRUE, main = "Smoothing")

tree.pred1 <- as.numeric(pred_tree)
ROC_tree <- roc(s_test$result, tree.pred1, percent = TRUE, main = "Smoothing")

rf.pred1 <- as.numeric(pred_rf$predictions)
ROC_rf <- roc(s_test$result, rf.pred1, percent = TRUE, main = "Smoothing")
```


**Logit AUC**
```{r}
auc(ROC_logit)
```


**Probit AUC**
```{r}
auc(ROC_probit)
```


**SVM AUC**
```{r}
auc(ROC_svm)
```


**GBM AUC**
```{r}
auc(ROC_gbm)
```


**Tree AUC**
```{r}
auc(ROC_tree)
```


**Random Forest AUC**
```{r}
auc(ROC_rf)
```


### plot ROC & AUC
```{r}
par(mar=c(1,1,1,1))  # create plotting environment
plot.roc(ROC_logit, s_test$result, percent = TRUE, main = "ROC curves", add =  FALSE, asp = NA)
lines(ROC_probit, col = "blue")
lines(ROC_svm, col = "red")
lines(ROC_gbm, col = "green")
lines(ROC_tree, col = "cyan")
lines(ROC_rf, col = "orange")
lines(ROC_svm_compare, col = "yellow")
axis(1, at = seq(0, 1, by=0.2), labels = paste(100*seq(1,0, by=-.2)), tick = TRUE)
legend(40, 20, 
       legend = c("model   AUC", "logit: 100", "probit: 100", "SVM: 100", "GBM: 100", "Tree: 100", "RF: 100", "Compare_SVM: 50"),
       col = c("white", "black", "blue", "red", "green", "cyan", "orange", "yellow"),
       lty = c(1, 1, 2, 3, 4, 5, 6, 6),
       pch = c(NA, NA, NA, NA, NA, NA, NA, NA),
       cex = 0.7)
```

